---
title: ''
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 2

For this assignment I'd like for you to explore the power of ggplot while also applying your data visualization critique knowledge to your new visualizations

### Step 1

Load tidyverse package. 
Load gapminder package and assign it to an object "data"

```{r data}
#your code below


```

### Step 2

Explore the variables and dimensions of your newly created object using the head() and dim() functions

```{r get to know your data, echo=FALSE}
#your code below


```

### Step 3

Create a data visualization that explores the relationship between year and life expectancy in the gapminder data

```{r data visualization, echo=FALSE}
#your code below


```

### Step 4
Evaluate your data visualization based on the 5 principles of an effective data visualizations we talked about in class on 9/4

---your evaluation here---

### Bonus Step 5 (earn an additional 5%)
Create a data visualization that add an additional variable to your visualization. A describe the resulting visualization (3 sentences or less)

```{r data visualization_bonus, echo=FALSE}
#your code below


```