---
title: "Class Notes"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment #3 Review

## Example 1

```{r}
#install.packages("socviz")
library(socviz)
library(tidyverse)
gs_data <- gss_sm
ggplot(data = gss_data, mapping = aes(x = bigregion)) +
  geom_bar(mapping = aes(y = ..prop.., group = 1))
```
## Example 2

```{r}
#install.packages("socviz")
ggplot(data = gss_data, mapping = aes(x = religion, fill = religion)) +
  geom_bar() +
  guides(fill = FALSE)
```
## Example 3

```{r}
bio <- read.csv("surveys_complete.csv", stringsAsFactors=FALSE)

ggplot(bio, aes(weight, hindfoot_length)) +
  geom_point(position = position_jitter())
```

## Example 4

```{r}
ggplot(subset(bio, species_id %in% c("DO", "DM", "DS") & sex %in% c("F", "M")),
       aes(x = sex, y = weight,  color = interaction(species_id,sex))) + facet_wrap( ~ species_id) +
       geom_point(alpha = 0.3, position = "jitter") +
       geom_boxplot(alpha = 0, colour = "black")
```

## TidyTuesday Workshop

#Step One - Load Data
```{r}

```

#Step Two - Explore Data
```{r}

```

#Step Three - Discuss as a group what kinds of visualizations you want to make

#Step Four - Sketch out your visualizations

#Step Five - Identify the building blocks of your visualizations

#Step Six - Write our your code - if you can't create your visualization identify what you want to create and why you can't create it with the knowledge you have of R. 

```{r}
library(tidyverse)
library(readr)
full_data <- read_csv("full_data.csv")


target <- c ("Rachel Green", "Ross Geller", "Joey Tribbiani", "Chandler Bing", "Monica Geller", "Phoebe Buffay")
fulldata_main <- filter (full_data, speaker %in% target)

p <- ggplot(data = subset(full_data, speaker %in% target), mapping = aes(x = emotion, fill = speaker)) +
  geom_bar()

p
